Supplementary Information 1

### 2025-04-01 index.Rmd
adding las  chgs in html

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      fig.path = "Figs/",
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

@@

## Overview

@@

```{r}
start1 <- SS_readstarter(file = file.path(dir1.1,
                                          "starter.ss"),
                              verbose = FALSE)
dat1 <- SS_readdat(file = file.path(dir1.2, start1$datfile),
                        verbose = FALSE)
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()
```

@@

```{r fig.width=9, fig.height=4}
dat1_std <- dat1$CPUE %>%
  group_by(index) %>%
  mutate(index_std = obs / mean(obs, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(index = recode(index,
                        `1` = "FISHERYBS",
                        `2` = "FISHERYEI",
                        `3` = "FISHERYGS",
                        `4` = "FISHERYJOIN",
                        `5` = "FISHERYSSIW",
                        `6` = "SURVEYBS",
                        `7` = "SURVEYEI",
                        `8` = "SURVEYGS",
                        `9` = "SURVEYJOIN",
                        `10` = "SURVEYSSIW",
                        `11` = "PREDATOR"))
ggplot(dat1_std, aes(x = year, y = index_std, group=index))+
  geom_point() +
  geom_smooth(method="lm",
              se=F)+
  facet_wrap(.~index,
             scales = "free_y",
             ncol=6)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  labs(title = "",
       x = "",
       y = "Standar index",
       color = "index")
```

```{r fig.width=10, fig.height=4}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

@@

# Expandir datos según la frecuencia
dat_expanded <- dat_long %>%
  uncount(weights = Frecuencia)

# Crear el gráfico con líneas verticales en cada facet
ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), 
                         fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1.8,
                               rel_min_height = 0.001,
                               col = "black",
                               fill = "grey") +
  facet_wrap(~FltSvy, ncol = 10) +
  geom_vline(data = mean_talla, aes(xintercept = mean_Talla), 
             color = "red", 
             size = 0.7,
             alpha=0.5) +
  labs(title = "",
       x = "cm",
       y = "",
       fill = "Densidad") +
  theme_few()
```

@@

## Run Models

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
directorios <- c("s1.1",
                 "s1.2",
                 "s1.3")

@@

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1.1,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

# s1.2
r4ss::run(
  dir = dir1.2,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

# s1.3
r4ss::run(
  dir = dir1.3,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)

# s1.4
r4ss::run(
  dir = dir1.4,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```

@@

Data used en both (spatial and No spatial models)

```{r }
SSplotData(base.model1.1, 
           subplots = 1,
           pheight = 15)

SSplotData(base.model1.4, 
           subplots = 1,
           pheight = 15)
```
# RESULTS

@@

### Main Variables poulation in `s1.1` scenario

@@

### Main Variables poulation in `s1.2` scenario

@@

### Main Variables poulation in `s1.3` scenario

@@

Selectivity

```{r fig.height=6}
SSplotSelex(base.model1.1,
            subplots = 1)

SSplotSelex(base.model1.2,
            subplots = 1)
```

```{r}
SSplotBiology(base.model1.1,
            subplots = 1:32)
```

```{r}
SSplotComps(base.model1.2,
            subplots = 1)

@@

```{r eval=FALSE}
total01 <- as.data.frame(base.model1.1$timeseries[1:31,c(2,5)]) %>%
  mutate(Serie = "No Spatial")
total1 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial implicit")
total2 <- as.data.frame(base.model2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ without S-R relation")
total3 <- as.data.frame(base.model3$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation weak (0.9 steepness)")
total4 <- as.data.frame(base.model4$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation strong (0.6 steepness)")
total5 <- as.data.frame(base.model5$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation mid-strong estimated ")
total6 <- as.data.frame(base.model6$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ Ricker S-R relation estimated")
total7 <- as.data.frame(base.model7$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ New set parameters Linf, k and M")

totalbio <- rbind(total01,
              total1,
              total2,
              total3,
              total4,
              total5,
              total6,
              total7)

ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
  geom_point(size=0.7,
             col="black") +
  geom_smooth(method = "loess",
              span=0.75, 
              se=TRUE)+
  labs(title = "Compare variables between scenarios",
       x = "",
       y = "Total biomass (mt)") +
  theme_few() +
  scale_color_viridis_d(option="H",
                        name="Scenarios")+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
```

```{r eval=FALSE}
SSplotComps(base.model01,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model1,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model2,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model3,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model4,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model5,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model6,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model7,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```
Heatmap

```{r}
df <- base.modeltest$lendbase

# Filtra el dataframe
df1 <- df %>% 
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  ))

@@

```{r eval=FALSE}
# Assuming your data is in a data frame called 'lendbase'
# Filter and select relevant columns
data <- base.model1$lendbase %>%
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  )) %>% 
  dplyr::select(Yr, Bin, Pearson)

@@

ggplot(data, aes(x = Yr, y = Bin)) +
  geom_hex(bins = 50) +
  scale_fill_gradient2(low = "blue", 
                       mid = "white", 
                       high = "red", midpoint = median(data$Bin)) +
  theme_minimal() +
  labs(title = "Hexagonal Heatmap of Pearson Residuals",
       x = "Year", y = "Bin", fill = "Pearson Residual")
```

@@

Step to do a good practice in model diagnosis is;

-     1.  Convergence. Final convergence criteria is 1.0e-04

-     2.  Residual (visual and metrics)

-     3.  Retrospective analysis (Mhon Parameter)

-    4.  Likelihood profile

all this framework try to follow recommendations of @Carvalho2021b

## Residual consistency

```{r}
SSplotRunstest(base.model1.1,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
SSplotRunstest(base.model1.2,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r eval=FALSE}
SSplotRunstest(base.model01,
               subplots = "len",
               add=T)
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
SSplotRunstest(base.model2,
               subplots = "len",
               add=T)
SSplotRunstest(base.model3,
               subplots = "len",
               add=T)
SSplotRunstest(base.model4,
               subplots = "len",
               add=T)
SSplotRunstest(base.model5,
               subplots = "len",
               add=T)
SSplotRunstest(base.model6,
               subplots = "len",
               add=T)
SSplotRunstest(base.modeltest,
               subplots = "len",
               add=T)
```

```{r}
SSplotRunstest(base.model1.1,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1.2,
               subplots = "cpue",
               add=T)
```

```{r eval=FALSE}
SSplotRunstest(base.model01,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model3,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model4,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model5,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model6,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.modeltest,
               subplots = "cpue",
               add=T)
```

```{r eval=FALSE}
SSplotJABBAres(base.model01,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1.2,
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model01,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1.2,
               subplots = "cpue",
               add=T)
```
## Relationship Stock-Recruit

```{r eval=FALSE}
l <- base.modeltest$recruit$SpawnBio
m <- base.modeltest$recruit$pred_recr
ggplot() +
  aes(l, m) +
  geom_point() +
  geom_smooth(method = "loess",
              se = TRUE, 
              color = "blue") +
  theme_few() +
  ylim(0, max(base.modeltest$recruit$pred_recr))

>>>>

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 6,
                      fig.width = 7,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      fig.path = "Figs/",
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

@@

# OVERVIEW

@@

```{r}
start1 <- SS_readstarter(file = file.path(dir1.4,
                                          "starter.ss"),
                              verbose = FALSE)
dat1 <- SS_readdat(file = file.path(dir1.4, start1$datfile),
                        verbose = FALSE)
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.4,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1.4, 
                                                "forecast.ss"),
                              verbose = FALSE)
# can also read in wtatage.ss for an empirical wt at age model using
# r4ss::SS_readwtatage()
```

@@

```{r fig.width=9, fig.height=4}
dat1_std <- dat1$CPUE %>%
  mutate(
    # Standardize 'obs' between 0 and 1
    index_std = (obs - min(obs)) / (max(obs) - min(obs)),

    # Recode 'index' values
    index = recode(index,
                   `1` = "FISHERYBS",
                   `2` = "FISHERYEI",
                   `3` = "FISHERYGS",
                   `4` = "FISHERYJOIN",
                   `5` = "FISHERYSSIW",
                   `6` = "SURVEYBS",
                   `7` = "SURVEYEI",
                   `8` = "SURVEYGS",
                   `9` = "SURVEYJOIN",
                   `10` = "SURVEYSSIW",
                   `11` = "PREDATOR"),

    # Convert 'index' to a factor with specific levels
    index = factor(index, levels = c("FISHERYBS", 
                                     "FISHERYEI", 
                                     "FISHERYGS",
                                     "FISHERYJOIN", 
                                     "FISHERYSSIW", 
                                     "SURVEYBS", 
                                     "SURVEYEI", 
                                     "SURVEYGS", 
                                     "SURVEYJOIN", 
                                     "SURVEYSSIW", 
                                     "PREDATOR")),

    # Assign categories based on 'index'
    category = case_when(
      index %in% c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW") ~ "Fishery",
      index %in% c("SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "SURVEYSSIW") ~ "Survey",
      index == "PREDATOR" ~ "Predator",
      TRUE ~ NA_character_  # In case there are any unexpected values
    )
  )

# Gráfico con 3 filas, dejando 'PREDATOR' en la última fila
ggplot(dat1_std, aes(x = year, y = index_std, group = index, color = category)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~index, scales = "free_y", ncol = 5) +  # Ajusta ncol para obtener 3 filas
  scale_color_manual(values = c("Fishery" = '#1b9e77', 
                                "Survey" = '#d95f02',
                                "Predator" = '#7570b3')) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "", x = "", y = "Standard index", color = "Category")

```{r fig.width=10, fig.height=4, warning=FALSE}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

@@

dat_expanded <- dat_long %>%
  uncount(weights = Frecuencia)

ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
  geom_density_ridges_gradient(scale = 1.8,
                               rel_min_height = 0.001,
                               col = "black",
                               alpha=0.4) +
  facet_wrap(~FltSvy, ncol = 10) +
  geom_vline(data = mean_talla, aes(xintercept = mean_Talla), 
             color = "red", 
             size = 0.7,
             alpha = 0.5) +
  scale_fill_manual(values = c("FISHERYBS" = '#1b9e77',  
                               "FISHERYEI" = '#1b9e77',  
                               "FISHERYGS" = '#1b9e77',
                               "FISHERYJOIN" = '#1b9e77',  
                               "FISHERYSSIW" = '#1b9e77',  
                               "SURVEYBS" = '#d95f02',  
                               "SURVEYEI" = '#d95f02',  
                               "SURVEYGS" = '#d95f02',  
                               "SURVEYJOIN" = '#d95f02',  
                               "SURVEYSSIW" = '#d95f02',  
                               "PREDATOR" = '#7570b3')) +  
  labs(title = "",
       x = "cm",
       y = "",
       fill = "Category") +
  theme_few()+
  theme(legend.position = "none")

@@

### Run Models

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
directorios <- c("s1.1",
                 "s1.2",
                 "s1.3",
                 "s1.4")

@@

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1.1,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)

# s1.2
r4ss::run(
  dir = dir1.2,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)

# s1.3
r4ss::run(
  dir = dir1.3,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)

# s1.4
r4ss::run(
  dir = dir1.4,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)
```

@@

Data used en both (spatial and No spatial models) `s1.1`

```{r }
SSplotData(base.model1.1, 
           subplots = 1,
           pheight = 15)
```

and `s1.4`

```{r }
SSplotData(base.model1.4, 
           subplots = 1,
           pheight = 15)
```

# RESULTS

@@

Main Variables poulation in `s1.1` scenario

@@

Main Variables poulation in `s1.2` scenario

@@

Main Variables poulation in `s1.3` scenario

@@

Main Variables poulation in `s1.4` scenario

```{r}
SSplotTimeseries(base.model1.4,
                 subplot = 1)
SSplotTimeseries(base.model1.4,
                 subplot = 9)
SSplotTimeseries(base.model1.4,
                 subplot = 7)
```
Selectivity

```{r}
SSplotSelex(base.model1.1,
            subplots = 1)
SSplotSelex(base.model1.2,
            subplots = 1)
SSplotSelex(base.model1.3,
            subplots = 1)
SSplotSelex(base.model1.4,
            subplots = 1)
```

```{r}
SSplotIndices(base.model1.1,
            subplots = 9)
SSplotIndices(base.model1.2,
            subplots = 9)
SSplotIndices(base.model1.3,
            subplots = 9)
SSplotIndices(base.model1.4,
            subplots = 9)
```

```{r}
SSplotComps(base.model1.1,
            subplots = 21)
SSplotComps(base.model1.2,
            subplots = 21)
SSplotComps(base.model1.3,
            subplots = 21)
SSplotComps(base.model1.4,
            subplots = 21)
```

@@

```{r}
total1 <- as.data.frame(base.model1.1$timeseries[,c(2,5)]) %>%
  mutate(Serie = "Without Ecosystem Variables")
total2 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "W/ Predator")
total3 <- as.data.frame(base.model1.3$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "W/ Environment")
total4 <- as.data.frame(base.model1.4$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "With Predator and Ecosystem")

totalbio <- rbind(total1,
              total2,
              total3,
              total4)

ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
  geom_point() +
  geom_smooth(method = "loess",
              span=0.5, 
              se=F)+
  labs(title = "",
       x = "",
       y = "Total biomass (mt)") +
  theme_few() +
  scale_color_viridis_d(option="H",
                        name="Scenarios")+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
```

Heatmap

```{r}
df <- base.model1.4$lendbase

df1 <- df %>% 
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  ))

@@

```{r eval=FALSE}
# Assuming your data is in a data frame called 'lendbase'
# Filter and select relevant columns
data <- base.model1.4$lendbase %>%
  dplyr::filter(Pearson<5) %>% 
  dplyr::select(c(1, 6, 19, 16)) %>% 
  dplyr::mutate(Fleet = dplyr::case_when(
    Fleet == 1 ~ "FISHERYBS",
    Fleet == 2 ~ "FISHERYEI",
    Fleet == 3 ~ "FISHERYGS",
    Fleet == 4 ~ "FISHERYJOIN",
    Fleet == 5 ~ "FISHERYSSIW",
    Fleet == 6 ~ "SURVEYBS",
    Fleet == 7 ~ "SURVEYEI",
    Fleet == 8 ~ "SURVEYGS",
    Fleet == 9 ~ "SURVEYJOIN",
    Fleet == 10 ~ "SURVEYSSIW",
    Fleet == 11 ~ "PREDATOR",
    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
  )) %>% 
  dplyr::select(Yr, Bin, Pearson)

@@

@@

A rigorous model diagnosis is essential to ensure the reliability and robustness of stock assessment models. The key steps for a good practice in model diagnosis include:

1. Convergence Check: The model must reach a final convergence criterion of 1.0e-04 to ensure numerical stability and reliable parameter estimation.

2. Residual Analysis: Both visual inspection and statistical metrics are used to evaluate model residuals, helping to detect patterns of bias or misfit.

3. Retrospective Analysis: The Mohn’s rho parameter is used to assess the consistency of model estimates when sequentially removing recent years of data, identifying potential overestimation or underestimation trends.

4. Likelihood Profile Analysis: This approach examines how the likelihood function behaves across a range of parameter values, providing insight into parameter uncertainty and model sensitivity.

This framework follows the recommendations outlined by @Carvalho2021b, aiming to enhance transparency and reproducibility in model evaluation.

### Residual consistency

Residual analysis is a critical component of model diagnostics in stock assessments. It helps evaluate the fit of the model to observed data and detect potential biases or inconsistencies. This process is applied to both length composition data and abundance indices such as CPUE (Catch Per Unit Effort) and survey-derived estimates.

For length composition data, residuals represent the difference between observed and model-predicted length distributions. The standardized residuals are calculated as the difference between observed and expected proportions at each length bin. These residuals are plotted by year to identify systematic trends, biases, or inconsistencies in the data. Ideally, they should be randomly distributed around zero, indicating no systematic over- or underestimation.

For abundance indices such as CPUE and fishery-independent surveys, residuals are analyzed to assess model fit and potential sources of bias. Residuals are computed as the difference between observed index values and those predicted by the model, typically standardized by dividing by the standard error to facilitate comparison across years. These residuals are then plotted over time to evaluate trends. A shaded confidence region, like the green area in the provided plot, represents expected variability, with outliers highlighted in red or other distinct markers. Persistent positive or negative residuals may indicate systematic bias in the model or data collection process.

Statistical diagnostics are also performed to check for autocorrelation in residuals, which can indicate potential model misspecifications. When mean residual values are close to zero, the model fit is considered unbiased.

By integrating these residual analyses for both length and abundance indices, stock assessment models can be refined, improving their reliability and increasing confidence in the assessment results.

```{r}
SSplotRunstest(base.model1.1,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
SSplotRunstest(base.model1.2,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
SSplotRunstest(base.model1.3,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
SSplotRunstest(base.model1.4,
               subplots = "len",
               add=T,
               plot = TRUE,
               plotdir = Figs)
```

```{r}
SSplotRunstest(base.model1.1,
               subplots = "cpue",
               add=T,
               print = TRUE)
SSplotRunstest(base.model1.2,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1.3,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1.4,
               subplots = "cpue",
               add=T)
```

### Residual Analysis and RMSE

Residual analysis of mean length data is a fundamental diagnostic tool in stock assessments. It helps evaluate whether the model provides an unbiased fit to the observed data and detects potential biases over time. In this figure, mean length residuals are plotted across years, differentiated by data source, including fishery-dependent (FISHERY) and fishery-independent (SURVEY) datasets, as well as predator-related observations (PREDATOR). The residuals represent the deviation of observed mean length from model-predicted values, standardized to facilitate interpretation.

The black line represents a locally estimated scatterplot smoothing (Loess) curve, which provides a trend line to visualize systematic deviations over time. The presence of persistent positive or negative trends in the residuals may indicate biases in the growth model, selectivity assumptions, or misrepresentation of recruitment variability. The gray bars highlight periods where residual variability is particularly high, suggesting potential inconsistencies between observed and predicted size structures.

RMSE quantifies the overall deviation between observed and predicted values, providing an aggregate measure of model fit. Lower RMSE values indicate better agreement between observed and predicted data. In fisheries stock assessment [@HurtadoF2015], RMSE thresholds for acceptable model performance typically range between *10% and 30%*, depending on the data quality and complexity of the population dynamics being modeled. Values exceeding this range suggest potential biases, requiring further investigation into the model structure, parameter estimation, or data sources.

By analyzing residual patterns and RMSE values, the model can be refined to improve the accuracy of mean length predictions, ultimately enhancing the reliability of stock assessment outcomes and management recommendations.

```{r}
SSplotJABBAres(base.model1.1,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1.2,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1.3,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1.4,
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model1.1,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1.2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1.3,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model1.4,
               subplots = "cpue",
               add=T)
```

### Comparision RMSE

```{r}
# Crear un data frame con los valores de RMSE
dfrmse <- data.frame(
  Model1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
  Model1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
  Model1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10]),
  Model1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])

)
summary(dfrmse)

# Comparación con boxplot
boxplot(dfrmse, main = "RMSE comparision", col = c('#ca0020','#f4a582','#bababa','#404040'))

# Prueba de diferencias estadísticas (ANOVA)
anova_result <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
summary(anova_result)

t.test(dfrmse$Model1.1, dfrmse$Model1.2)  # Comparar Model1.1 vs Model1.2
t.test(dfrmse$Model1.1, dfrmse$Model1.3)  # Comparar Model1.1 vs Model1.3
t.test(dfrmse$Model1.2, dfrmse$Model1.3)
t.test(dfrmse$Model1.3, dfrmse$Model1.4)
# ANOVA para ver si hay diferencias globales
anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
summary(anova_rmse)

# Prueba post-hoc si ANOVA es significativa
#TukeyHSD(anova_rmse)

dfrmse_long <- reshape2::melt(dfrmse)

ggplot(dfrmse_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(width=0.3) +
  geom_jitter(width = 0.01, alpha = 0.3) +
  theme_minimal() +
  scale_fill_manual(name= "Scenario",
                    values = c("Model1.1" = "#ca0020", 
                               "Model1.2" = "#f4a582", 
                               "Model1.3" = "#bababa", 
                               "Model1.4" = "#404040")) +
  labs(title = "", x = "", y = "RMSE")

```
This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4) used to evaluate recruitment estimates of Antarctic krill. RMSE serves as an indicator of model accuracy, with lower values representing better predictive performance.

s1.1 exhibits the lowest median RMSE, suggesting it has the best overall fit among the four models. In contrast, Models 1.2, 1.3, and 1.4 show higher RMSE values, indicating comparatively lower predictive accuracy. The interquartile range (IQR) of these models is relatively similar, suggesting comparable variability in RMSE across models. Additionally, s1.1 has an outlier above 1.5 RMSE, which could indicate a case where the model's predictions deviated significantly from observed values.

Overall, this analysis highlights that incorporating different environmental or predator-related variables in the models impacts their predictive ability. The differences in RMSE suggest that some models may overfit or underfit the recruitment patterns of krill, emphasizing the need to refine model selection based on ecological and statistical considerations.

### Relationship Stock-Recruit

\[
R = \frac{aS}{1 + bS}
\]

Where:  
- \( S \) is the spawning stock biomass.  
- \( R \) is the predicted recruitment.  
- \( a \) is the maximum recruitment capacity.  
- \( b \) regulates density dependence (higher values of \( b \) result in a lower plateau).

The blue line will show the asymptotic curve that describes the relationship between spawning stock biomass and recruitment.

```{r}
# First dataset
l <- base.model1.1$recruit$SpawnBio  
m <- base.model1.1$recruit$pred_recr 
yr <- base.model1.1$timeseries$Yr[1:35]

# Second dataset
leco <- base.model1.4$recruit$SpawnBio  
meco <- base.model1.4$recruit$pred_recr 
yreco <- base.model1.4$timeseries$Yr[1:35]

# Beverton-Holt models
bev_holt_model1 <- nls(m ~ (a * l) / (1 + b * l), 
                        start = list(a = max(m), b = 1 / max(l)), 
                        data = data.frame(l, m))
